---
title: Fact Checker DRAFT
subtitle: Fact Checker of BMs of IPBES Assessments
date: today
author:
  - name: Rainer M Krug 
    id: rmk
    orcid: 1234-1234-1234-1234
    email: Rainer@krugs.de
    affiliation: 
      - name: University of Zürich
        city: Zürich
        state: ZH
        url: www.uzh.ch
    roles: [author, editor]
abstract: > 
  The assessments Linked Open Data is used to conduct analysis and searches 
  to discover new knowledge related to the assessments.
# keywords:
#   - aaaaa
#   - bbbbb
license: "CC BY"
citation: 
  type: report
  doi: xxxxxxxxxxxxxx
doi: xxxxxxxxxxxxxx4
version: 0.0.1

format:
    html:
        toc: true
        toc-depth: 6
        toc_expand: true
        embed-resources: false
        code-fold: true
        code-summary: 'Show the code'
        keep-md: false
        # grid:
        #     sidebar-width: 0px
        #     body-width: 4000px
        #     margin-width: 200px
        #     gutter-width: 1.5rem    
---

```{r}
#| label: setup
#| include: false

library(openalexR)
library(dplyr)
library(arrow)
library(ggplot2)

library(IPBES.R)
library(htmlwidgets)
library(htmltools)

library(tictoc)

# Self-contained-report helper: copy heavy standalone HTML artifacts (DT
# tables, plotly explorers, rendered TD_*.qmd design docs) into the report's
# own _files/ sidecar directory so distributing the report only requires
# IPBES_Fact_Checker.html + IPBES_Fact_Checker_files/ — not a separate
# output/tables/ directory (or the TD_*.html files) alongside it. Returns
# the report-relative link/iframe-src path(s).
#
# Deliberately file.path("output/reports", ...) rather than a bare
# knitr::current_input()-derived name: the qmd source lives in
# input/reports/, but the rendered .html itself gets moved to
# output/reports/ by the _targets.R target AFTER this chunk runs (Quarto
# renders next to the input regardless of execute_dir) -- writing the
# sidecar directly into its final home means it's already a correct
# sibling of the .html once that move happens, no second relocation step
# needed for this directory.
report_files_dir <- file.path(
  "output/reports",
  paste0(tools::file_path_sans_ext(knitr::current_input()), "_files")
)
dir.create(report_files_dir, recursive = TRUE, showWarnings = FALSE)
copy_into_report_files <- function(paths) {
  file.copy(paths, report_files_dir, overwrite = TRUE)
  file.path(basename(report_files_dir), basename(paths))
}
```

# Introduction
IPBES assessments were converted / enhanced into [Linked Open Data (LOD)](https://github.com/IPBES-Data/IPBES_LOD) using the ttl format. 

Three assessments available in ttl format are currently:

- the [First Global Assessment](https://github.com/IPBES-Data/IPBES_LOD/blob/main/Global%20Assessment%201/README_GA1.md) (GA1)
- the [Invasive Alien Species Assessment](https://github.com/IPBES-Data/IPBES_LOD/tree/main/Invasive%20Alien%20Species%20Assessment) (IAS).
<!-- - the [Values Assessment](https://github.com/IPBES-Data/IPBES_LOD/tree/main/Values%20Assessment) (VA) -->

# Methods

```{r}
#| label: td_doc_paths
#| include: false

# Rendered by the td_doc_html target (see _targets.R) from the TD_<name>.qmd
# design docs (each this report's format/layout block followed by its own
# prose) and copied into the report's own _files/ sidecar so the links
# below work in a distributed copy of the report too.
td_doc_html_paths <- copy_into_report_files(targets::tar_read(td_doc_html))
td_doc_link <- function(name) {
  td_doc_html_paths[grepl(paste0("/", name, "\\.html$"), td_doc_html_paths)]
}
```

Methodology is documented separately in the `TD_*` design documents. **Note:
these are currently stale relative to the implementation** — treat them as a
record of the design rationale rather than an up-to-date specification of the
pipeline as it stands today.

- [TD_targets](`r td_doc_link("TD_targets")`) — the `targets` pipeline architecture
- [TD_BM_NLI_approach](`r td_doc_link("TD_BM_NLI_approach")`) — the active NLI scoring approach
- [TD_LLM_approach](`r td_doc_link("TD_LLM_approach")`) — the earlier LLM-comparison approach (**superseded, source removed**)
- [TD_NLI_LLM_two_phase](`r td_doc_link("TD_NLI_LLM_two_phase")`) — the two-phase NLI → LLM design (**implemented**; not yet wired into this report's tables/figures)
- [TD_NLI_training](`r td_doc_link("TD_NLI_training")`) — NLI fine-tuning plan (**planned, not yet implemented**)
- [TD_formatting](`r td_doc_link("TD_formatting")`) — archived report design system

The figures and tables in the Results section below are all produced by
`targets` pipeline targets — see `_targets.R` and the corresponding
`R/build_*.R` files for the exact computation behind each one.

# Results

```{r}
#| label: results_data
#| include: false

refs_paths <- targets::tar_read(refs_parquet)
works_paths <- targets::tar_read(works_parquet)
works_citing_paths <- targets::tar_read(works_citing_parquet)

refs <- dplyr::bind_rows(lapply(refs_paths, function(p) {
  arrow::open_dataset(p) |> dplyr::select(doi) |> dplyr::collect()
}))

key_papers <- dplyr::bind_rows(lapply(works_paths, function(p) {
  arrow::open_dataset(p) |>
    dplyr::select(km, bm, doi, publication_year, type) |>
    dplyr::collect()
})) |>
  dplyr::mutate(oa_input = TRUE)

citing_papers <- dplyr::bind_rows(lapply(works_citing_paths, function(p) {
  arrow::open_dataset(p) |>
    dplyr::select(km, bm, doi, publication_year, type) |>
    dplyr::collect()
})) |>
  dplyr::mutate(oa_input = FALSE)

all_papers <- dplyr::bind_rows(key_papers, citing_papers)

# Figure and table artifact paths — resolved via tar_read() so the report
# always shows what the pipeline actually last built, never a stale/hardcoded
# path.
fig_pub_per_year_paths <- copy_into_report_files(targets::tar_read(fig_pub_per_year))
fig_pub_per_year_png <- fig_pub_per_year_paths[grepl(
  "\\.png$",
  fig_pub_per_year_paths
)]
fig_pub_per_year_pdf <- fig_pub_per_year_paths[grepl(
  "\\.pdf$",
  fig_pub_per_year_paths
)]
fig_pub_per_year_svg <- fig_pub_per_year_paths[grepl(
  "\\.svg$",
  fig_pub_per_year_paths
)]

overlap_key_paper_html <- targets::tar_read(overlap_key_paper_table)
overlap_key_paper_html <- overlap_key_paper_html[grepl(
  "\\.html$",
  overlap_key_paper_html
)]
overlap_key_paper_html <- copy_into_report_files(overlap_key_paper_html)

overlap_after_2018_sub_messages_html <- targets::tar_read(
  overlap_after_2018_sub_messages_table
)
overlap_after_2018_sub_messages_html <- overlap_after_2018_sub_messages_html[grepl(
  "\\.html$",
  overlap_after_2018_sub_messages_html
)]
overlap_after_2018_sub_messages_html <- copy_into_report_files(
  overlap_after_2018_sub_messages_html
)

overlap_after_2018_background_messages_html <- targets::tar_read(
  overlap_after_2018_background_messages_table
)
overlap_after_2018_background_messages_html <- overlap_after_2018_background_messages_html[grepl(
  "\\.html$",
  overlap_after_2018_background_messages_html
)]
overlap_after_2018_background_messages_html <- copy_into_report_files(
  overlap_after_2018_background_messages_html
)
```

The dataset contains **`r refs |> dplyr::distinct(doi) |> nrow()`** papers
referenced in the background messages. Out of these, **`r key_papers |> dplyr::distinct(doi) |> nrow()`**
were usable for the snowball search and also yielded results.

```{r}
#| label: numbers_table
#|

all_papers |>
  dplyr::summarise(
    n = dplyr::n(),
    n_key_paper = sum(oa_input, na.rm = TRUE),
    n_non_key_paper = sum(!oa_input, na.rm = TRUE),
    n_nkp_pre_2018 = sum(
      (!oa_input) & (publication_year <= 2018),
      na.rm = TRUE
    ), # Count of citing papers with publication_year <= 2018
    n_nkp_post_2018 = sum(
      (!oa_input) & (publication_year > 2018),
      na.rm = TRUE
    ), # Count of citing papers with publication_year > 2018
    .by = c(km, bm)
  ) |>
  dplyr::arrange(km, bm) |>
  IPBES.R::table_dt()
```

## Publication types after 2018

Here we show the different types of publication after 2018 as identified by the snowbal search. All Sub-background messages are grouped.

```{r}
#| label: publication_types_grouped
#| echo: false

all_papers |>
  dplyr::filter(
    publication_year > 2018
  ) |>
  dplyr::summarize(
    n = dplyr::n(),
    .by = type
  ) |>
  dplyr::arrange(
    dplyr::desc(n)
  ) |>
  knitr::kable()
```

This table shows the number of work types per sub-background message. It allows filtering as well as sorting and downloading.

```{r}
#| label: publication_types_non_grouped
#| echo: false

all_papers |>
  dplyr::filter(
    publication_year > 2018
  ) |>
  dplyr::summarize(
    n = dplyr::n(),
    .by = c(type, km, bm)
  ) |>
  dplyr::arrange(
    dplyr::desc(n)
  ) |>
  IPBES.R::table_dt(
    fixed_columns = NULL
  )
```

## Publications per year

![](`r fig_pub_per_year_png`)

- [Download `png`](`r fig_pub_per_year_png`)
- [Download `pdf`](`r fig_pub_per_year_pdf`)
- [Download `svg`](`r fig_pub_per_year_svg`)

## Overlap

In this section we will look at the number of background messages in which a given paper, as identified by their OpenAlex id. We willlook at that separatr for key-papers and non key-papers.

In the following tables only the papers which occur in more than one background message are included and only papers which were published after 2018.

This overlap can have different causes. Either there is an overlap of the key-papers, ehich results in an overlap of the new publications, or there is a topical overlap betweent the new paper themselves. 

### Overlap of Key Papers in Background Messages

The first table shows the overlap of the key-papers only regarding background messages.

Only papers occurring in more than 1 background message are shown.

<iframe src="`r overlap_key_paper_html`" width="100%" height="500px"></iframe>

### Overlap of Papers after 2018 in Sub Messages

The second table shows the overlap of all papers after 2018 regarding sub messages.

Only papers with in more then 5 Sub Messages are shown.

<iframe src="`r overlap_after_2018_sub_messages_html`" width="100%" height="500px"></iframe>

### Overlap of Papers after 2018 in Background Messages

The third table shows the overlap of all papers after 2018 regarding background messages.

Only papers with in more then 5 Background Messages are shown.

<iframe src="`r overlap_after_2018_background_messages_html`" width="100%" height="500px"></iframe>

## NLI Alignment Scores

Each citing work is scored against its partition's Background Message (BM) as `SUPPORTS`, `REFUTES`, or `NOT_ENOUGH_INFO` using a zero-shot NLI model. See [TD_BM_NLI_approach](`r td_doc_link("TD_BM_NLI_approach")`) for methodology.

```{r}
#| label: nli_overview_paths
#| include: false

nli_overview_data_paths <- targets::tar_read(nli_overview_data)
nli_overview_figures_paths <- copy_into_report_files(targets::tar_read(nli_overview_figures))
# copy_into_report_files() (defined in the results_data chunk above) copies
# the self-contained widgets into IPBES_Fact_Checker_files/ and returns the
# report-relative iframe src paths directly.
nli_bm_explorer_paths <- copy_into_report_files(targets::tar_read(
  nli_bm_explorer_html
))
```

```{r}
#| label: nli_overview_sections
#| results: asis

for (data_path in nli_overview_data_paths) {
  x <- readRDS(data_path)
  nli_active_this <- x$nli_active %||% "deberta_zeroshot"
  gran_this <- x$granularity %||% "naive_bm"
  cat(sprintf("\n### %s (%s, %s)\n\n", x$assessment, nli_active_this, gran_this))

  if (isTRUE(x$empty)) {
    cat("No NLI scores available yet for this assessment/model/granularity combination.\n\n")
    next
  }

  # Exact suffix match (not a loose substring grepl) -- filenames now carry
  # both a model suffix ("" for deberta_zeroshot, "_<nli_active>" otherwise)
  # and a granularity suffix (always "_<granularity>", explicit for every
  # value including naive_bm), so a bare "_<assessment>\\.png$" anchor
  # would silently only ever match one branch and miss the others.
  model_suffix_this <- if (identical(nli_active_this, "deberta_zeroshot")) "" else paste0("_", nli_active_this)
  gran_suffix_this <- paste0("_", gran_this)
  suffix_this <- paste0(model_suffix_this, gran_suffix_this)

  cat(sprintf(
    "- **%s** works scored\n- **%d** background messages\n- **%d** key messages\n- **%s%%** uncertain (confidence < 0.60)\n\n",
    format(x$n_total, big.mark = ","),
    x$n_bm,
    x$n_km,
    x$pct_unc
  ))

  fig_this <- nli_overview_figures_paths[
    grepl(paste0("_", x$assessment, suffix_this, "\\.png$"), nli_overview_figures_paths)
  ]
  plot_overall <- fig_this[grepl("plot_overall", fig_this)]
  plot_km <- fig_this[grepl("plot_km", fig_this)]
  plot_bm <- fig_this[grepl("plot_bm", fig_this)]
  plot_conf <- fig_this[grepl("plot_conf", fig_this)]
  plot_aln <- fig_this[grepl("plot_aln", fig_this)]

  cat("#### Label distribution\n\n")
  cat(
    knitr::kable(
      x$label_overall |>
        dplyr::mutate(
          label = factor(
            label,
            levels = c("SUPPORTS", "NOT_ENOUGH_INFO", "REFUTES")
          )
        ) |>
        dplyr::arrange(label),
      col.names = c("Label", "Works", "%")
    ),
    sep = "\n"
  )
  cat("\n\n")
  if (length(plot_overall)) {
    cat(sprintf("![](%s)\n\n", plot_overall))
  }

  cat("#### Interactive BM Explorer\n\n")
  cat(
    "Select a background message from the dropdown and a minimum confidence ",
    "from the slider. All works are always shown; the slider instead splits ",
    "each bar into a **solid** segment (confidence at or above the ",
    "threshold) and a **hollow/outlined** segment on top (below the ",
    "threshold, stacked) — in the label-distribution panel (each bar also ",
    "labelled `n / N`: works at/above the threshold vs. all works with that ",
    "label), the confidence panel (with a dashed line marking the ",
    "threshold), and the alignment panel. Mean/median confidence and the % ",
    "meeting the threshold are reported above the panels. Below the panels, ",
    "a second dropdown (default **REFUTES**) selects which label the ",
    "drill-down table lists — the actual works for the current BM + ",
    "threshold + label, sorted by confidence descending and capped at 50 ",
    "rows (the caption reports the true total when capped). Each work links ",
    "to its DOI where one exists (OpenAlex otherwise). **Download table ",
    "(CSV)** is not limited to the 50 rows shown — it exports up to 5,000 ",
    "matching works for the current BM + threshold + label (sorted by ",
    "confidence descending, flagged with a trailing comment line if even ",
    "that limit truncates the true match count), including the assessment ",
    "and BM as explicit columns.\n\n"
  )
  explorer_this <- nli_bm_explorer_paths[
    grepl(paste0("_", x$assessment, suffix_this, "\\.html$"), nli_bm_explorer_paths)
  ]
  if (length(explorer_this)) {
    cat(sprintf(
      '<iframe src="%s" width="100%%" height="1020px" style="border:none;"></iframe>\n\n',
      explorer_this
    ))
  }

  cat("#### Per Key Message\n\n")
  cat(
    knitr::kable(
      x$label_km,
      col.names = c("KM", "SUPPORTS %", "NEI %", "REFUTES %")
    ),
    sep = "\n"
  )
  cat("\n\n")
  if (length(plot_km)) {
    cat(sprintf("![](%s)\n\n", plot_km))
  }

  cat("#### Per Background Message\n\n")
  if (length(plot_bm)) {
    cat(sprintf("![](%s)\n\n", plot_bm))
  }
  cat(
    knitr::kable(
      x$table_bm,
      col.names = c("KM", "BM", "SUPPORTS %", "NEI %", "REFUTES %")
    ),
    sep = "\n"
  )
  cat("\n\n")

  cat("#### Confidence by label\n\n")
  cat(
    knitr::kable(
      x$conf_table,
      col.names = c("Label", "n", "Mean", "Median", "P25", "P75")
    ),
    sep = "\n"
  )
  cat("\n\n")
  if (length(plot_conf)) {
    cat(sprintf("![](%s)\n\n", plot_conf))
  }

  cat("#### Alignment score per BM\n\n")
  cat(
    knitr::kable(
      x$align_table,
      col.names = c(
        "KM",
        "BM",
        "n",
        "Mean aln",
        "SD",
        "SUPP %",
        "REF %",
        "NEI %",
        "Unc %"
      )
    ),
    sep = "\n"
  )
  cat("\n\n")
  if (length(plot_aln)) cat(sprintf("![](%s)\n\n", plot_aln))

  # AI Based Interpretation — last subsection of each assessment. Hand-
  # written commentary, deliberately scoped to the two signals of interest
  # (SUPPORTS and REFUTES) rather than the dominant NOT_ENOUGH_INFO label.
  # Recomputed against the current scored data (below) — will go stale as
  # scoring continues or the NLI config changes, so re-derive the numbers
  # from x$table_bm / x$label_km before trusting this after a rerun.
  cat("#### AI Based Interpretation\n\n")
  if (identical(x$assessment, "GA1")) {
    cat(r"----(
::: {.callout-note}
## About this section
Focused specifically on the SUPPORTS/REFUTES signals — NOT_ENOUGH_INFO (~71% of GA1 pairs) is set aside here as the least informative label. Written against the current scored data (2026-07-06); re-derive the numbers below from the tables above before trusting this after a rerun.
:::

**Key finding — KM D is the only reliable signal.** GA1 overall SUPPORTS (16.1%) edges out REFUTES (13.0%), but that gap is almost entirely driven by KM D (34.7% SUPPORTS vs 10.9% REFUTES). KMs A–C sit close to parity or worse — A: 15.0% vs 12.5%, B: 10.0% vs 14.2% (REFUTES-inflated), C: 14.5% vs 20.9% (REFUTES-inflated) — so the argmax label there carries little information on its own.

Within KM D, four BMs are unambiguous: **D3** (62.6% SUPPORTS vs 4.4% REFUTES), **D2** (44.0% vs 6.5%), **D1** (40.8% vs 10.3%), and **D12** (40.7% vs 11.1%) — all over 40% SUPPORTS at single-digit-to-low-teens REFUTES. The one exception inside KM D is **D10**, where REFUTES (15.2%) exceeds SUPPORTS (11.6%) — an outlier against its own KM's pattern, worth treating differently from its D-series siblings.

Likely reason: D-series BMs are probably more concrete and empirically testable — biodiversity metrics, species trends, ecosystem services — the kind of claims abstracts engage with directly.

::: {.callout-warning}
## REFUTES inflation — KMs A, B, C

The BMs with the widest REFUTES-over-SUPPORTS gap:

| BM | SUPPORTS | REFUTES | Gap (REFUTES − SUPPORTS) |
|----|---------|---------|---------|
| C6  |  5.2 % | 28.6 % | +23.4 |
| A9  | 13.8 % | 30.7 % | +16.9 |
| C8  |  8.3 % | 24.8 % | +16.5 |
| B11 | 15.4 % | 28.4 % | +13.0 |
| C7  |  7.1 % | 16.1 % |  +9.0 |
| B3  |  3.3 % | 12.0 % |  +8.7 |

Citing papers of an assessment being scored as majority-REFUTES at these margins is not credible. This is the well-documented zero-shot NLI failure mode: the hypothesis *"This paper refutes the following claim: …"* draws spurious entailment when the BM contains complex, multi-part, or politically nuanced language. The model fires on surface-level topical tension rather than genuine contradiction.
:::

::: {.callout-note}
## Closest to parity — weakest signal

A handful of BMs have SUPPORTS and REFUTES within a couple of points of each other — the argmax label there is close to a coin flip:

- **A1** 12.9 % vs 11.0 %
- **B10** 12.7 % vs 14.7 %
- **A4** 9.7 % vs 11.8 %

For these, only the continuous alignment score (`p_supports − p_refutes`) carries any real information.
:::

##### Notable exception in KM A

**A2** (20.5% SUPPORTS, 11.0% REFUTES) is the clearest positive signal in KMs A–C outside of KM D — a nearly 10-point SUPPORTS lead, well ahead of its KM-A neighbours. Its claim is likely more concrete or empirically specific than the rest of KM A — worth inspecting manually as a calibration reference.

##### Operational implications

1. **Do not act on raw REFUTES labels for KMs A–C** without Phase-2 LLM review. NLI here is a coarse NEI filter, not a judgment.
2. **KM D SUPPORTS can be treated as the most reliable signal in the assessment** — D3, D2, D1, D12 in particular (all >40% SUPPORTS, single-digit-to-low-teens REFUTES). D10 is the one KM D BM that does NOT fit this pattern.
3. **Use `p_supports − p_refutes` for ranking** rather than argmax labels, especially for the near-parity BMs above.
4. **`max_length: 256` is probably hurting KMs A–C.** Truncating abstract tails removes findings — the part that matters most for support/refute judgements on conceptual BMs. The base-model-at-512 comparison (`deberta_zeroshot_v3`) is the highest-priority next experiment.
)----")
    cat(sprintf(
      "5. **Fine-tuning on BM citations** ([TD_NLI_training](%s)) would most help KMs A–C — the support/refute axis that zero-shot is fumbling. D-series BMs are already working.\n\n",
      td_doc_link("TD_NLI_training")
    ))
  } else if (identical(x$assessment, "IAS")) {
    cat(r"----(
::: {.callout-note}
## About this section
Focused specifically on the SUPPORTS/REFUTES signals — NOT_ENOUGH_INFO (~81% of IAS pairs) is set aside here as the least informative label. Written against the current scored data (2026-07-06); re-derive the numbers below from the tables above before trusting this after a rerun.
:::

**Key finding — IAS is weaker and more REFUTES-skewed than GA1.** Overall, IAS REFUTES (11.1%) actually *exceeds* SUPPORTS (8.4%) — the opposite of GA1's overall SUPPORTS lean (16.1% vs 13.0%). Unlike GA1, no IAS BM reaches D-series levels of confidence; the strongest SUPPORTS signal is markedly weaker.

::: {.callout-tip}
## Strongest SUPPORTS signal

**D33** (25.9% SUPPORTS vs 9.5% REFUTES, a 16-point lead) is the clearest positive signal in IAS, appearing under both KM-D4 and KM-D7. A handful of others lean the same way but at much lower absolute rates: **B13** (19.4% vs 10.8%), **D30** (8.7% vs 3.0%), **D27** (7.5% vs 1.8%), and **D25** (12.5% vs 7.2%). None of these approach GA1's D3/D2/D1/D12 margins.
:::

::: {.callout-warning}
## REFUTES inflation

The BMs with the widest REFUTES-over-SUPPORTS gap — the same zero-shot failure mode documented for GA1's KMs A–C:

| BM | SUPPORTS | REFUTES | Gap (REFUTES − SUPPORTS) |
|----|---------|---------|---------|
| A1  |  7.5 % | 22.6 % | +15.1 |
| A8  |  3.6 % | 16.7 % | +13.1 |
| D32 | 22.5 % | 33.5 % | +11.0 |
| A4  |  7.6 % | 17.6 % | +10.0 |
| C19 |  5.0 % | 14.3 % |  +9.3 |
| D28 |  8.1 % | 14.7 % |  +6.6 |

**D32** is a special case: both SUPPORTS and REFUTES are elevated (22.5% and 33.5%) rather than REFUTES alone — it appears under three different KMs (KM-D5, KM-D6, KM-D7), which may mean it's a genuinely contested or multi-faceted claim rather than a pure model artifact, and is worth a manual look before writing it off as noise.
:::

##### Operational implications

1. **Do not act on raw REFUTES labels anywhere in IAS** without Phase-2 LLM review — the overall REFUTES-over-SUPPORTS skew makes this an even higher priority here than for GA1's KMs A–C.
2. **D33 is the one IAS BM currently worth treating as a real signal** — everything else is either near parity or REFUTES-leaning.
3. **Use `p_supports − p_refutes` for ranking** rather than argmax labels; IAS's overall skew means a plain SUPPORTS-label count will systematically undercount real support.
4. **D32 deserves manual review before either accepting or discarding it** — high confidence in both directions, across multiple KMs, is a different failure mode from the single-direction REFUTES inflation seen elsewhere.
)----")
  } else {
    cat("No AI-based interpretation has been written yet for this assessment.\n\n")
  }
}
```

# Funnel Reports

Companion per-assessment reports narrow REFUTES- and SUPPORTS-labelled
citing works down through NLI certainty and LLM re-verification. See each
report's own Methodology section for the full 3-level sieve definition.

```{r}
#| label: label_funnel_links
#| results: asis

# Plain relative links, not iframes/copy_into_report_files(): each funnel
# report is a full standalone page meant to be opened on its own, and both
# it and this report sit as flat siblings both in the project root and in
# output/reports/ (build_report_output_dir()), so a relative link resolves
# in both places with no copying needed here.
label_funnel_html_paths <- c(
  targets::tar_read(report_refutes_funnel_html),
  targets::tar_read(report_supports_funnel_html)
)
# Assessment ids are alphanumeric (GA1, IAS); the trailing blob is
# "_<model>" (optional -- omitted only for the deberta_zeroshot default)
# and "_<granularity>" (always present -- every granularity, including
# naive_bm, is explicit now). Both segments can contain underscores
# themselves, so the combined trailing blob is split by checking whether it
# ENDS in one of the fixed, known granularity values (same list
# _targets.R's own nli_granularities uses) rather than a naive underscore
# split, which would be ambiguous for e.g. "bge_m3_zeroshot_complete_bm".
known_granularities <- c("naive_bm", "complete_bm", "atomic_bm")
for (p in label_funnel_html_paths) {
  m <- regmatches(
    basename(p),
    regexec("^IPBES_([A-Z]+)_Report_([A-Za-z0-9]+)(?:_(.+))?\\.html$", basename(p))
  )[[1]]
  label <- m[[2]]
  id <- m[[3]]
  suffix <- m[[4]]
  gran <- "naive_bm"
  model <- "deberta_zeroshot"
  if (nzchar(suffix)) {
    hit <- known_granularities[vapply(known_granularities, function(g) grepl(paste0("(^|_)", g, "$"), suffix), logical(1))]
    if (length(hit)) {
      gran <- hit[[1]]
      suffix <- sub(paste0("_?", gran, "$"), "", suffix)
    }
    if (nzchar(suffix)) model <- suffix
  }
  cat(sprintf("- [%s %s Funnel Report (%s, %s)](%s)\n", id, label, model, gran, basename(p)))
}
```

# QA Reports

Per-assessment QA check for how `nli_ready_evidence_parquet` actually
split each Background Message into claims under the currently active
`granularity` setting — not a scoring result. See the report's own
Methodology section for details.

```{r}
#| label: bm_split_report_links
#| results: asis

# Same plain-relative-link convention as the Funnel Reports section above.
bm_split_html_paths <- targets::tar_read(bm_split_report_html)
for (p in bm_split_html_paths) {
  m <- regmatches(
    basename(p),
    regexec("^QA_BM_Split_Report_([A-Za-z0-9]+)(?:_(.+))?\\.html$", basename(p))
  )[[1]]
  id <- m[[2]]
  gran <- if (nzchar(m[[3]])) m[[3]] else "naive_bm"
  cat(sprintf("- [%s QA BM Split Report (%s)](%s)\n", id, gran, basename(p)))
}
```

Per-(assessment, granularity) QA check on Phase 1 (zero-shot NLI) scoring
itself — claims x citing works with title/abstract/DOI and label/
confidence, plus a label x confidence-decile matrix. Not a scoring result;
see the report's own Methodology section for details.

```{r}
#| label: nli_scores_qa_report_links
#| results: asis

# Same suffix-splitting logic as the Funnel Reports section above --
# model_suffix and gran_suffix are both optional and concatenated in the
# filename, so known_granularities (defined there) disambiguates them here
# too.
nli_scores_qa_html_paths <- targets::tar_read(nli_scores_qa_report_html)
for (p in nli_scores_qa_html_paths) {
  m <- regmatches(
    basename(p),
    regexec("^QA_NLI_Scores_Report_([A-Za-z0-9]+)(?:_(.+))?\\.html$", basename(p))
  )[[1]]
  id <- m[[2]]
  suffix <- m[[3]]
  gran <- "naive_bm"
  model <- "deberta_zeroshot"
  if (nzchar(suffix)) {
    hit <- known_granularities[vapply(known_granularities, function(g) grepl(paste0("(^|_)", g, "$"), suffix), logical(1))]
    if (length(hit)) {
      gran <- hit[[1]]
      suffix <- sub(paste0("_?", gran, "$"), "", suffix)
    }
    if (nzchar(suffix)) model <- suffix
  }
  cat(sprintf("- [%s QA NLI Scores Report (%s, %s)](%s)\n", id, model, gran, basename(p)))
}
```

Per-(assessment, llm_config) QA check on Phase 2 (LLM verification) —
independent LLM review of NLI-flagged claim/work pairs, plus a full audit
of every key paper regardless of its own NLI label/confidence. Not a
scoring result; see the report's own Methodology section for details.

```{r}
#| label: llm_verification_qa_report_links
#| results: asis

# Filename is always "..._<assessment>_<llm_config>.html" -- llm_config is
# always present (no optional-default-unsuffixed convention here, unlike
# nli_model_suffix()'s historical exception for "deberta_zeroshot").
llm_verification_qa_html_paths <- targets::tar_read(llm_verification_qa_report_html)
for (p in llm_verification_qa_html_paths) {
  m <- regmatches(
    basename(p),
    regexec("^QA_LLM_Verification_Report_([A-Za-z0-9]+)_(.+)\\.html$", basename(p))
  )[[1]]
  id <- m[[2]]
  cfg <- m[[3]]
  cat(sprintf("- [%s QA LLM Verification Report (%s)](%s)\n", id, cfg, basename(p)))
}
```

