---
title: QA LLM Verification Report
subtitle: QA check for Phase 2 (LLM verification) of NLI-flagged claim/work pairs
date: today
params:
  assessment_id: "GA1"
  llm_config: "openrouter_cheap"
format:
    html:
        toc: true
        toc-depth: 4
        embed-resources: true
        code-fold: true
        code-summary: 'Show the code'
        keep-md: false
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(DT)

# No report_files_dir/copy_into_report_files() here, and no source()ing of
# R/build_llm_verification_qa_data.R either -- quarto::quarto_render() runs
# this file in its own fresh session that never sources R/*.R (same
# reasoning QA_NLI_Scores_Report.qmd's own setup chunk gives). The DT
# widgets were already built and cached by build_llm_verification_qa_data()
# -- library(DT) above is what lets the cached htmlwidget objects render
# correctly once reloaded.
expected_name <- function(ext) {
  sprintf("llm_verification_qa_%s_%s.%s", params$assessment_id, params$llm_config, ext)
}

data_path <- targets::tar_read_raw("llm_verification_qa_data") |>
  (\(x) x[basename(x) == expected_name("rds")])()
x <- readRDS(data_path)

# Embedded as base64 data: URIs (xfun::base64_uri()), same convention
# QA_NLI_Scores_Report.qmd already uses and already confirmed necessary --
# Quarto/pandoc's own resource resolution mangles plain paths here even
# when absolute.
decile_path <- targets::tar_read_raw("llm_verification_qa_figures") |>
  (\(paths) paths[basename(paths) == sprintf("llm_verification_qa_decile_%s_%s.png", params$assessment_id, params$llm_config)])()
alluvial_path <- targets::tar_read_raw("llm_verification_qa_figures") |>
  (\(paths) paths[basename(paths) == sprintf("llm_verification_qa_alluvial_%s_%s.png", params$assessment_id, params$llm_config)])()
```

# `r params$assessment_id` -- llm_config: `r params$llm_config`

## Methodology

This report is a **QA check on Phase 2 (LLM verification), not a scoring
result in its own right** — it lets each LLM-reviewed claim/work pair be
spot-checked by eye (title, abstract, DOI, NLI's own label/confidence, the
LLM's independent label/agreement/quote), shows how the two classifiers'
labels compare, and — separately — audits **every** key paper the keypaper
LLM chain has reviewed, regardless of NLI's own label or confidence. See
`TD_NLI_LLM_two_phase.qmd` for the full Phase 2 design and
`QA_NLI_Scores_Report.qmd` for the equivalent Phase 1 (NLI) QA report.

A literal ternary probability plot (Phase 1's own QA figure) was considered
and rejected for Phase 2: the LLM emits one categorical verdict, not a
3-part probability composition, so a simplex density would really just be
NLI's own chart wearing a Phase-2 label. The two figures below are built
instead from what Phase 2 actually produces.

```{r}
#| label: empty-check
#| results: asis
if (isTRUE(x$empty)) {
  cat("No LLM-verified data available yet for this assessment/config combination — Phase 2 hasn't produced output here.\n\n")
  knitr::knit_exit()
}
```

```{r}
#| label: summary
#| results: asis
cat(sprintf(
  "- **%s** LLM-reviewed (claim, citing work) pairs in total\n",
  format(x$n_total, big.mark = ",")
))
cat(sprintf(
  "- **%.1f%%** of reviewed pairs: `llm_label` agrees with NLI's own `label`\n",
  x$llm_agrees_pct
))
cat(sprintf(
  "- **%.1f%%** of reviewed pairs: the LLM found `sufficient_evidence` to make a call\n",
  x$sufficient_evidence_pct
))
cat(sprintf(
  "- **%.1f%%** of reviewed pairs: demoted to NOT_ENOUGH_INFO because the LLM's cited quote wasn't found verbatim in the supplied text\n",
  x$quote_verbatim_demotion_pct
))
cat(sprintf(
  "- **%.1f%%** of reviewed pairs: `direct_evidence_match` (the citing work traces back to a seed reference matching the claim's own evidence braces)\n",
  x$direct_evidence_match_pct
))
cat(sprintf(
  "- The table below shows up to **%d** highest-NLI-confidence rows per claim (**%s** rows shown here, out of the %s total) — every claim's own true row count is carried in the `n_total_claim` column so no truncation is silent\n",
  x$per_claim_cap, format(x$n_shown, big.mark = ","), format(x$n_total, big.mark = ",")
))
```

## NLI label x LLM label

A literal confusion matrix comparing the two classifiers on the same
pairs — Phase 2 has no numeric confidence of its own to bucket by, unlike
Phase 1's label x confidence-decile matrix.

```{r}
#| label: confusion-matrix
knitr::kable(x$confusion_matrix)
```

## Does LLM/NLI agreement track NLI's own confidence?

X-axis: NLI's own confidence, binned into deciles. Y-axis: the % of
LLM-reviewed pairs in that decile where `llm_label` agrees with NLI's
`label` — one line, built on the full reviewed set (point size = how many
pairs support that decile). The natural expectation: agreement should rise
as NLI's own confidence rises.

The orange overlay is the **key-paper LLM validation chain**
(`llm_verification_keypaper_parquet`) — every key/seed paper reviewed,
regardless of its own NLI label or confidence: individual jittered points
at each key paper's own NLI confidence (100 if the LLM agreed with NLI for
that paper, 0 if not), plus a single dashed reference line at their
aggregate agreement rate, directly comparable to the general-population
line above it since both use the same `llm_agrees` metric. Absent (general
line only) until that chain has been run for this assessment.

```{r}
#| label: decile-figure
#| results: asis
if (length(decile_path)) {
  cat(sprintf('<img src="%s" style="max-width:100%%;"><br><br>\n\n', xfun::base64_uri(decile_path)))
}
```

## NLI to LLM label flow

Band width = pair count flowing from each NLI label to each LLM label (the
exact same 3x3 contingency as the table above) — coloured green where the
two classifiers agree, orange where they don't. The proportional picture
(how much of REFUTES's mass actually stays LLM-confirmed REFUTES vs. gets
downgraded or flipped) reads faster from a flow diagram than from the raw
counts above.

The right-hand panel is the same flow diagram restricted to key papers only
(its own, separate scale — a key-paper flow plotted on the main panel's
scale would be visually invisible given how much smaller that corpus is).
Absent (main panel only) until the key-paper LLM verification chain has
been run.

```{r}
#| label: alluvial-figure
#| results: asis
if (length(alluvial_path)) {
  cat(sprintf('<img src="%s" style="max-width:100%%;"><br><br>\n\n', xfun::base64_uri(alluvial_path)))
}
```

## Key-paper LLM validation

Every key/seed paper IPBES cites as evidence for a BM, reviewed by the LLM
**regardless of its own NLI label or confidence** — unlike the main
citing-works corpus, which is routed by `nli_labels`/`nli_certainty` for
cost control, key papers are a small, bounded, high-importance set where
every one is worth an independent check. A key paper *is* the evidence a
BM was written from, so any LLM verdict other than SUPPORTS deserves a
human look.

```{r}
#| label: keypaper-summary
#| results: asis
if (is.null(x$keypaper_n)) {
  cat("No key-paper LLM verification data available yet for this assessment/config — that chain hasn't been run.\n\n")
} else {
  cat(sprintf(
    "- **%d** key papers reviewed (all of them)\n- **%.1f%%** LLM-confirmed as SUPPORTS (the expected outcome)\n- **%.1f%%** where the LLM agrees with NLI's own label\n\n",
    x$keypaper_n, x$keypaper_llm_supports_pct, x$keypaper_llm_agrees_pct
  ))
  if (!is.null(x$keypaper_flagged_widget)) {
    cat("Key papers the LLM did **not** confirm as SUPPORTS — the actionable subset:\n\n")
  } else {
    cat("Every key paper was LLM-confirmed as SUPPORTS.\n\n")
  }
}
```

```{r}
#| label: keypaper-flagged-table
if (!is.null(x$keypaper_flagged_widget)) {
  x$keypaper_flagged_widget
}
```

## Claims and scores

Use the filter row below the header to narrow by KM, BM, claim, or label.

```{r}
#| label: scores-table
x$widget
```
