---
title: QA NLI Training Data Report
subtitle: QA check for the NLI fine-tuning training data export
date: today
params:
  assessment_id: "GA1"
  nli_config: "bge_m3_zeroshot_atomic_bm"
  granularity: "atomic_bm"
format:
    html:
        toc: true
        toc-depth: 4
        embed-resources: true
        code-fold: true
        code-summary: 'Show the code'
        keep-md: false
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(DT)

# No report_files_dir/copy_into_report_files() here, and no source()ing of
# R/build_nli_training_qa_data.R either -- quarto::quarto_render() runs this
# file in its own fresh session that never sources R/*.R (same reasoning
# every other QA_*_Report.qmd in this project already documents). The DT
# widget was already built and cached by build_nli_training_qa_data() --
# library(DT) above is what lets the cached htmlwidget object render
# correctly once reloaded.
expected_name <- function(ext) {
  sprintf(
    "nli_training_qa_%s_%s_%s.%s",
    params$assessment_id, params$nli_config, params$granularity, ext
  )
}

data_path <- targets::tar_read_raw("nli_training_qa_data") |>
  (\(x) x[basename(x) == expected_name("rds")])()
x <- readRDS(data_path)
```

# `r params$assessment_id` -- nli_config: `r params$nli_config` -- granularity: `r params$granularity`

## Methodology

This report is a **QA check on the NLI fine-tuning training data export
(`R/build_nli_training_data.R`), not a scoring result in its own right** —
it lets every exported `(premise, hypothesis, label)` pair be spot-checked
by eye before spending time fine-tuning on it. See `TD_NLI_training.qmd`
for the full training-data strategy this export implements.

Positives are key papers the LLM confirmed `SUPPORTS` with a real verbatim
quote — not every cited paper unconditionally, since only a small fraction
of key papers actually have explicit abstract-level support for their own
claim. Negatives are real NLI/LLM disagreements (NLI called `SUPPORTS` or
`REFUTES` confidently, but the LLM's grounded, quote-checked review said
`NOT_ENOUGH_INFO`), downsampled to roughly match the positive count.
`REFUTES` examples are the real `nli_label == llm_label == "REFUTES"` rows,
included unconditionally.

```{r}
#| label: empty-check
#| results: asis
if (isTRUE(x$empty)) {
  cat("No training data available yet for this assessment/config/granularity combination — the export hasn't produced output here.\n\n")
  knitr::knit_exit()
}
```

```{r}
#| label: summary
#| results: asis
cat(sprintf("- **%d** training pairs exported in total\n", x$n_total))
cat("- Class balance:\n")
for (lab in names(x$label_counts)) {
  cat(sprintf("  - `%s`: %d\n", lab, x$label_counts[[lab]]))
}
cat("- By source:\n")
for (src in names(x$source_counts)) {
  cat(sprintf("  - `%s`: %d\n", src, x$source_counts[[src]]))
}
cat("- Key paper vs. citing-work premise:\n")
for (kp in names(x$keypaper_counts)) {
  cat(sprintf("  - `keypaper=%s`: %d\n", kp, x$keypaper_counts[[kp]]))
}
```

## Training pairs

Use the filter row below the header to narrow by label, source, key-paper origin, KM, or BM.

```{r}
#| label: pairs-table
x$widget
```
