---
title: QA NLI Scores Report
subtitle: QA check for Phase 1 (zero-shot NLI) scoring of claims against citing works
date: today
params:
  assessment_id: "GA1"
  granularity: "naive_bm"
  nli_config: "deberta_zeroshot"
format:
    html:
        toc: true
        toc-depth: 4
        embed-resources: true
        code-fold: true
        code-summary: 'Show the code'
        keep-md: false
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(DT)

# No report_files_dir/copy_into_report_files() here, and no source()ing of
# R/build_nli_scores_qa_data.R either -- quarto::quarto_render() runs this
# file in its own fresh session that never sources R/*.R (same reasoning
# IPBES_Label_Funnel_Report.qmd's setup chunk gives for inlining
# gran_suffix()/nli_model_suffix() rather than sourcing R/branch_helpers.R
# there). The DT widget itself was already built and cached by
# build_nli_scores_qa_data(), so this chunk only needs to read it back and
# print it -- library(DT) above is what lets the cached htmlwidget object
# render correctly once reloaded.
gran_suffix <- paste0("_", params$granularity)
model_suffix <- if (identical(params$nli_config, "deberta_zeroshot")) "" else paste0("_", params$nli_config)

expected_name <- function(ext) {
  sprintf(
    "nli_scores_qa_%s%s%s.%s",
    params$assessment_id, model_suffix, gran_suffix, ext
  )
}

data_path <- targets::tar_read_raw("nli_scores_qa_data") |>
  (\(x) x[basename(x) == expected_name("rds")])()
x <- readRDS(data_path)

# Embedded as a base64 data URI (via xfun::base64_uri(), a knitr/xfun
# dependency already present) rather than through a report_files_dir/
# copy_into_report_files() sidecar (unlike IPBES_Label_Funnel_Report.qmd's
# figures) or a plain markdown image path -- confirmed directly that even
# an absolute path handed to `![](...)` gets mangled by Quarto/pandoc's own
# resource resolution here (it silently turned "/Users/.../x.png" into
# "./Users/.../x.png" and then warned the resource couldn't be fetched,
# rather than erroring loudly). Encoding to a data: URI ourselves sidesteps
# that path-resolution step entirely, and needs no sidecar directory for a
# single PNG.
#
# Matches build_nli_scores_qa_figures.R's own sprintf() pattern exactly
# ("nli_scores_qa_ternary_<id><suffixes>.png").
expected_ternary_name <- function() {
  sprintf(
    "nli_scores_qa_ternary_%s%s%s.png",
    params$assessment_id, model_suffix, gran_suffix
  )
}
ternary_path <- targets::tar_read_raw("nli_scores_qa_figures") |>
  (\(x) x[basename(x) == expected_ternary_name()])()
```

# `r params$assessment_id` -- granularity: `r params$granularity` -- nli_config: `r params$nli_config`

## Methodology

This report is a **QA check on Phase 1 (zero-shot NLI) scoring, not a
scoring result in its own right** — it lets each claim's scored citing
works be spot-checked by eye (title, abstract, DOI, label, confidence, and
the underlying class probabilities), and shows the overall
SUPPORTS/REFUTES/NOT_ENOUGH_INFO x confidence distribution. It covers
**Phase 1 only** — see `QA_LLM_Verification_Report.qmd` for the equivalent
Phase 2 (LLM verification) QA report. See `TD_BM_NLI_approach.qmd` for the
full NLI approach and `TD_NLI_LLM_two_phase.qmd` for how Phase 2 builds on
this.

```{r}
#| label: empty-check
#| results: asis
if (isTRUE(x$empty)) {
  cat("No scored data available yet for this assessment/granularity/config combination — Phase 1 hasn't produced output here.\n\n")
  knitr::knit_exit()
}
```

```{r}
#| label: summary
#| results: asis
cat(sprintf(
  "- **%s** scored (claim, citing work) pairs in total\n",
  format(x$n_total, big.mark = ",")
))
cat(sprintf(
  "- The table below shows up to **%d** highest-confidence rows per claim (**%s** rows shown here, out of the %s total) — every claim's own true row count is carried in the `n_total_claim` column so no truncation is silent; use the filters or the confusion matrix above for the full picture\n",
  x$per_claim_cap, format(x$n_shown, big.mark = ","), format(x$n_total, big.mark = ",")
))
```

## Label x confidence matrix

REFUTES/SUPPORTS/NOT_ENOUGH_INFO counts across confidence deciles, computed
on the **full** scored dataset (not the capped table below).

```{r}
#| label: confusion-matrix
knitr::kable(x$matrix)
```

## Class probability distribution (ternary)

Since `p_supports + p_refutes + p_nei = 1` always, every row's three class
probabilities live on a 2D triangular surface rather than filling a 3D
volume — this is the standard way to show exactly that kind of 3-part
composition losslessly in 2D. Each corner is 100% one label; a point's
position encodes all three probabilities at once, and its color is the
local **density** of scored pairs at that combination (log-scaled so
structure is visible everywhere, not just at the tallest peaks). The three
solid, coloured lines are the actual decision boundaries — where the
winning label flips from one to another — running from the centroid out to
each edge's midpoint; each corner's percentage is the real share of rows
that label actually won.

The black dashed triangle is the **certain/uncertain boundary**: `label` is
always the argmax of the three probabilities (which line above a point
falls on decides that), but `confidence` — the *winning* probability — is a
separate question, and a row is `uncertain` whenever `confidence` is below
this granularity's own `uncertain_threshold` (shown in the legend). Points
inside the dashed triangle are `uncertain` regardless of which label
technically won there; only the three small corners beyond it are
`certain`. A label can be assigned with real confidence right up against
this line while still counting as uncertain — the two are independent
judgements, not the same thing read two ways.

When available, small white markers overlay the actual **key papers** —
the seed/reference works IPBES cites as evidence for a BM, scored against
that same BM's claim (averaged across whatever claims exist under that
BM — we don't yet have claim-level evidence-to-paper mapping, only
BM-level) — as a QA sanity check: a key paper *is* the evidence a BM was
written from, so it should overwhelmingly land in the SUPPORTS region; if
it doesn't, that's worth investigating, not a normal finding. Each
corner's bracketed percentage is the key papers' own share for that label,
next to the full-corpus share, so the two can be compared directly, followed
by the p-value from a two-proportion test (key-paper share vs. full-corpus
share for that corner) — a small p-value means the key-paper share is
genuinely different from the full-corpus one, not just sampling noise from
the much smaller key-paper count. This overlay comes from a separate
scoring pass (`nli_scores_keypaper_evidence`) that now runs automatically
alongside the main citing-works scoring — it's absent here only on a cold
first build, before that chain has produced anything yet.

```{r}
#| label: ternary-figure
#| results: asis
if (length(ternary_path)) {
  cat(sprintf('<img src="%s" style="max-width:100%%;"><br><br>\n\n', xfun::base64_uri(ternary_path)))
}
if (!is.null(x$keypaper_summary)) {
  r <- x$keypaper_summary
  cat(sprintf(
    "**Key paper validation:** of **%d** key papers, **%.1f%%** landed in SUPPORTS\n\n",
    r$n, r$pct_supports
  ))
}
```

## Claims and scores

Use the filter row below the header to narrow by KM, BM, claim, or label.

```{r}
#| label: scores-table
x$widget
```
