---
title: QA NLI Fine-Tuned Model Report
subtitle: QA check for a locally fine-tuned NLI classifier (scripts/training/train_nli.py)
date: today
params:
  nli_config: "bge_m3_zeroshot_atomic_bm"
format:
    html:
        toc: true
        toc-depth: 4
        embed-resources: true
        code-fold: true
        code-summary: 'Show the code'
        keep-md: false
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(DT)
library(ggplot2)

# No source()ing of R/build_nli_finetuned_model_qa_data.R here --
# quarto::quarto_render() runs this file in its own fresh session that
# never sources R/*.R (same reasoning every other QA_*_Report.qmd in this
# project already documents). Everything displayable was already built and
# cached by build_nli_finetuned_model_qa_data() -- this qmd just reads it
# back and lays it out.
expected_name <- function(ext) {
  sprintf("nli_finetuned_model_qa_%s.%s", params$nli_config, ext)
}

data_path <- targets::tar_read_raw("nli_finetuned_model_qa_data") |>
  (\(x) x[basename(x) == expected_name("rds")])()
x <- readRDS(data_path)
```

# nli_config: `r params$nli_config`

## Methodology

This report is a **QA check on a locally fine-tuned NLI classifier
(`scripts/training/train_nli.py`), not a scoring result in its own right**
— it lets a training run's real composition, class balance, and held-out
performance be spot-checked before deciding whether the resulting model is
worth deploying. See `TD_NLI_training.qmd` for the full training-data
strategy and fine-tuning design this implements.

Only produced when the active nli config has `train: true` set in
`input/config.yaml` — otherwise `nli_finetuned_model` writes an empty
marker and this report shows the empty state below rather than a real
result.

```{r}
#| label: empty-check
#| results: asis
if (isTRUE(x$empty)) {
  cat(sprintf(
    "No fine-tuned model for `nli_config: %s` — either `train: false` in input/config.yaml (the default), or a run that failed before completing. Set `train: true` for this config and re-run `tar_make(names = \"nli_finetuned_model\")` to produce one.\n\n",
    params$nli_config
  ))
  knitr::knit_exit()
}
```

```{r}
#| label: summary
#| results: asis
cat(sprintf("- **Run**: `%s` (%s)\n", x$run_id, x$timestamp))
cat(sprintf("- **Base model**: `%s`\n", x$model_id))
cat(sprintf("- **Premise mode**: `%s`\n", x$premise_mode))
cat(sprintf(
  "- **Assessments**: %s — **Granularity**: %s\n",
  paste(x$assessments, collapse = ", "), paste(x$granularity, collapse = ", ")
))
cat(sprintf(
  "- **%d** total rows (**%d** train / **%d** held-out eval)\n",
  x$n_rows_total, x$n_rows_train, x$n_rows_eval
))
cat("- Class balance:\n")
for (lab in names(x$label_counts)) {
  cat(sprintf("  - `%s`: %d\n", lab, x$label_counts[[lab]]))
}
cat("- Key paper vs. citing-work premise:\n")
for (kp in names(x$keypaper_counts)) {
  cat(sprintf("  - `keypaper=%s`: %d\n", kp, x$keypaper_counts[[kp]]))
}
cat(sprintf(
  "- **Held-out accuracy: %.1f%%** (best checkpoint: `%s`, eval_loss = %.4f)\n",
  100 * x$accuracy, basename(x$best_model_checkpoint), x$best_metric
))
```

## Held-out classification report

Computed on the `r x$n_rows_eval` rows this run never trained on. `support`
is the number of true examples of that class in the held-out split — with
this few rows, treat individual precision/recall values as indicative, not
precise.

```{r}
#| label: classification-table
DT::datatable(
  x$classification_df, rownames = FALSE, options = list(dom = "t", paging = FALSE)
)
```

## Training curve

Eval loss at the end of each epoch — should trend down across epochs if
the model is genuinely learning (a flat line at `log(3) ≈ 1.099` — the
loss of a classifier guessing uniformly across 3 classes — is the signature
of a training run that never actually moved away from its random head
initialization).

```{r}
#| label: loss-curve
if (!is.null(x$loss_plot)) {
  x$loss_plot
} else {
  cat("No eval_loss entries in the training log.")
}
```

Full per-step log (loss, grad_norm, learning_rate where logged; eval_loss
at each epoch boundary):

```{r}
#| label: log-table
DT::datatable(
  x$log_df, rownames = FALSE,
  options = list(dom = "t", paging = FALSE, scrollX = TRUE)
)
```
